\huge Open, privacy-preserving protocols for lawful surveillance

### 2016-02-28 paper/intro.tex
...
fleshed out intro and biblio

>>>>

Continuing the approach taken by Segal {\it et al.}~\cite{sff-foci2014}, we
seek to design and implement protocols for {\it accountable surveillance}.
We require that government surveillance be conducted according to {\it open
processes}, as defined in Section~\ref{sec-open} below, and that the privacy
of {\it untargeted users} be protected.  We consider the surveillance goals
of {\it set intersection} and {\it contact chaining} and show that both can 
be achieved in a privacy-preserving, accountable fashion.

The utility of set-intersection protocols was demonstrated in the 
High Country Bandits case~\cite{anderson13cell}.
After obtaining cell-tower dumps -- sets of about 150,000 total users
whose cell phones had been in the vicinity of three banks at the times
that those banks were robbed -- the FBI intersected the sets and discovered 
that a single phone had been used at all of the relevant times in all of the 
relevant places.  They arrested the owner
of that phone and were able to prove that he was one of the robbers.
Although this FBI dragnet was effective in catching robbers, it also swept
in the cell-phone numbers of approximately 149,999 innocent bystanders.
Segal {\it et al.}~\cite{sff-foci2014} provide an accountable protocol for
set intersection that preserves the privacy of innocent bystanders.  Their
rudimentary implementation requires just under two hours on a test instance
with 150,000 total users.  In Section~\ref{sec-intersection} below, we provide
a more careful implementation that is faster by a factor of 10; in particular,
it runs for approximately 10.5 minutes on the test instance of 150,000 users.

In Section~\ref{sec-chaining}, we turn our attention to 
accountable-surveillance protocols for contact chaining.

### 2016-02-28 paper/open.tex
fleshed out intro and biblio

\section{The Openness Principle in Lawful Surveillance}
In this section, we review the openness principle put forth by Segal 
{\it et al.}~\cite{sff-foci2014}.  Readers familiar with \cite{sff-foci2014} 
may skip to the next section.

@@

More precisely, Segal {\it et al.}~\cite{sff-foci2014} draw a distinction 
between two classes of communication-system users.  {\it Targeted users} are 
those who are under suspicion and are targets of properly authorized warrants.
All others are {\it untargeted users}; they are the vast majority of all 
users of a general-purpose, mass-communication system.  Segal {\it et 
al.}~\cite{sff-foci2014} posit that the
following {\it Openness Principle} should govern all surveillance activity
in a democratic society:

\begin{enumerate}
\item[I]
Any surveillance or law-enforcement process
that obtains or uses private information
about untargeted users shall be
an open, public, unclassified process.
\item[II]
Any secret surveillance or law-enforcement processes shall use only:
\begin{enumerate}
\item[(a)] public information, and
\item[(b)] private information about targeted users obtained under 
authorized warrants via open surveillance processes.
\end{enumerate}
\end{enumerate}
Segal {\it et al.} interpret this principle as a requirement that an open
``privacy firewall'' be placed between government agencies and citizens'
private information in a mass-communication network.  Processes that move
untargeted users' private information through the firewall must be open
processes.  More generally, the openness principle asserts that, by using
appropriate security technology, government agencies can make their 
data-collection {\it processes} fully public without revealing sensitive
{\it content} of a specific investigation.  For a more detailed explanation,
see Section 2 of \cite{sff-foci2014}.

>>>>

\section{The Openness Principle in Lawful Surveillance}\label{sec-open}
In this section, we review the openness principle put forth by Segal 
{\it et al.}~\cite{sff-foci2014}.  Readers familiar with \cite{sff-foci2014} 
may skip to the next section.

@@

More precisely, Segal {\it et al.}~\cite{sff-foci2014} draw a distinction 
between two classes of communication-system users.  {\it Targeted users} are 
those who are under suspicion and are targets of properly authorized warrants.
All others are {\it untargeted users}; they are the vast majority of all 
users of a general-purpose, mass-communication system.  Segal {\it et 
al.}~\cite{sff-foci2014} posit that the
following {\it Openness Principle} should govern all surveillance activity
in a democratic society:
\begin{enumerate}
\item[I]
Any surveillance or law-enforcement process
that obtains or uses private information
about untargeted users shall be
an open, public, unclassified process.
\item[II]
Any secret surveillance or law-enforcement processes shall use only:
\begin{enumerate}
\item[(a)] public information, and
\item[(b)] private information about targeted users obtained under 
authorized warrants via open surveillance processes.
\end{enumerate}
\end{enumerate}
Segal {\it et al.} interpret this principle as a requirement that an open
``privacy firewall'' be placed between government agencies and citizens'
private information in a mass-communication network.  Processes that move
untargeted users' private information through the firewall must be open
processes.  The targeted class contains both {\it known} users, {\it i.e.},
those for whom the government has a name, address, phone number, email address,
or other piece of personally identifying information, and {\it unknown} users.
It is not, as it may seem on the surface, oxymoronic to call a user both
``targeted'' and ``unknown,'' because ambient information may justify the
targeting of an individual without identifying him or her in any standard sense
of ``identify.''  For example, a government agency may obtain a ``John Doe
warrant''~\cite{bieber-penn2002} to investigate users who were present in
locations $L_i$ at times $T_i$, for $1\leq i\leq k$, without being able to 
identify those users, because relevant events occurred at those locations at
those times. Segal {\it et al.}~\cite{sff-foci2014} show how an 
accountable-surveillance protocol can be used to obtain a large set of
{\it encrypted} data about both targeted and untargeted users, feed it into a 
cryptographic protocol that winnows it down to the records of users targeted by
the John Doe warrant, and decrypt {\it only} those records.  Thus, targeted
unknown users can be identified ({\it i.e.}, turned into targeted, known users)
without government agencies' identifying any untargeted users whose 
encrypted records are touched by the surveillance process.

The essence of the openness principle is that, by using
appropriate security technology, government agencies can make their 
data-collection {\it processes} fully public without revealing sensitive
{\it content} of a specific investigation.  For a more detailed explanation,
see Section 2 of \cite{sff-foci2014}.
